Computing solution space properties of combinatorial optimization problems via generic tensor networks

### 2022-09-02 paper/paper.tex
Updates from Overleaf

\begin{lemma}
    The number of the largest $k$ independent sets of a graph $G = (V, E)$ can be computed in time $O(K\log K)\cc$.
\end{lemma}
The proof is similar to proving \Cref{thm:complexpoly}, except only $k$ largest coefficients are involved in the addition and multiplication.

@@

Similarly, we have the following lemma by combining the sum-product expression tree algebra with the truncated polynomial.

\begin{lemma}\label{thm:complexexprk}
        Let $G = (V, E)$ be a graph and $\mathcal{I}_k$ be the number of independent sets with size $k$. The independent sets with size $k > \alpha(G)-K$ of can be computed in time $O(K\log K)\cc + O(|V|\sum_{k=\alpha(G)-K+1}^{\alpha(G)}|\mathcal{I}_k|)$.
\end{lemma}

@@

Similarly, if one is only interested in obtaining independent sets with largest $K$ sizes, we have the following lemma.
\begin{lemma}
    Let $G = (V, E)$ be a graph, its independent sets with size $k > \alpha(G)-K$ of can be unbiasedly sampled in time $O(K\log K)\cc + O(|V|^2 M)$, where $M$ is the number of samples.
\end{lemma}

@@

\begin{theorem}
    Let $G = (V, E, W)$ be a weighted graph, its $K$ largest independent sets sizes can be computed in time $O(K\log K)\cc$.
\end{theorem}
\begin{proof}
$k$ largest independent sets can be computed by contracting a tensor network with extended tropical numbers. The vertex tensor and edge tensor are
\begin{equation}
    \left(W^{{\rm T}k}\right)^{(v)} = \left(\begin{matrix}
        \mymathbb{1} \\
        -\infty^{\otimes k-1} \otimes w_v
    \end{matrix}\right),   
    \qquad
        \left(B^{{\rm T}k}\right)^{(u, v)} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right),
\end{equation}
where we have used $x_v^{w_v} = (-\infty^{\otimes k-1} \otimes 1)^{w_v} = -\infty^{\otimes k-1} \otimes w_v$.

@@

\begin{lemma}
    Let $G = (V, E, W)$ be a weighted graph, its $K$ largest independent sets can be computed in time $O(K\log K|V|)\cc$.
\end{lemma}
To find solutions corresponding to the largest $k$ sizes, one can combine the extended tropical algebra with the bit string algebra (\Cref{eq:singleconfig}).
Since the $\oplus$ operation of the configuration sampler will not be used in the combined algebra, the resulting configurations are deterministic and complete.

>>>>

\begin{theorem}
    The number of the largest $k$ independent sets of a graph $G = (V, E)$ can be computed in time $O(K\log K)\cc$.
\end{theorem}
The proof is similar to proving \Cref{thm:complexpoly}, except only $k$ largest coefficients are involved in the addition and multiplication.

@@

Similarly, we have the following theorem by combining the sum-product expression tree algebra with the truncated polynomial.

\begin{theorem}\label{thm:complexexprk}
        Let $G = (V, E)$ be a graph and $\mathcal{I}_k$ be the number of independent sets with size $k$. The independent sets with size $k > \alpha(G)-K$ of can be computed in time $O(K\log K)\cc + O(|V|\sum_{k=\alpha(G)-K+1}^{\alpha(G)}|\mathcal{I}_k|)$.
\end{theorem}

@@

Similarly, if one is only interested in obtaining independent sets with largest $K$ sizes, we have the following theorem.
\begin{theorem}
    Let $G = (V, E)$ be a graph, its independent sets with size $k > \alpha(G)-K$ of can be unbiasedly sampled in time $O(K\log K)\cc + O(|V|^2 M)$, where $M$ is the number of samples.
\end{theorem}

@@

\begin{theorem}
    Let $G = (V, E)$ be a weighted graph, its $K$ largest independent sets sizes can be computed in time $O(K\log K)\cc$.
\end{theorem}
\begin{proof}
$k$ largest independent sets can be computed by contracting a tensor network with extended tropical numbers. The vertex tensor and edge tensor are
\begin{equation}
    \left(W^{{\rm T}k}\right)^{(v)} = \left(\begin{matrix}
        \mymathbb{1} \\
        -\infty^{\otimes k-1} \otimes w_v
    \end{matrix}\right),   
    \qquad
        \left(B^{{\rm T}k}\right)^{(u, v)} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right),
\end{equation}
where we have used $x_v^{w_v} = (-\infty^{\otimes k-1} \otimes 1)^{w_v} = -\infty^{\otimes k-1} \otimes w_v$.

@@

\begin{theorem}
    Let $G = (V, E)$ be a weighted graph, its $K$ largest independent sets can be computed in time $O(K\log K|V|)\cc$.
\end{theorem}
To find solutions corresponding to the largest $k$ sizes, one can combine the extended tropical algebra with the bit string algebra (\Cref{eq:singleconfig}).
Since the $\oplus$ operation of the configuration sampler will not be used in the combined algebra, the resulting configurations are deterministic and complete.
